---
title: "Penetrance Probability Distributions"
author: "Brett"
date: "January 04, 2018"
output: 
  pdf_document: 
    highlight: pygments
    toc: true
    toc_depth: 4
---

#Introduction: 
This file is used to visualize probability distributions derived from the SCN5A dataset. several representative variants' probability distributions are plotted. At the end, probability distributions are calculated for representative "classified variants" using the classification scheme put forward by the ACMG. The final calculations are the probabilities variants classifiable as (likely) pathogenic, VUS, or (likley) benign have a true penetrance (individuals presenting with either BrS1 or LQT3) greater than 20%. There is also a sensitivity analysis done at the end.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rms)
library(pROC)
library(tableone)
library(ggplot2)
library("RSQLite")
```

#Read in data

```{r data, fig.width=5}
con = dbConnect(SQLite(), 
dbname="/Users/B/Dropbox/SCN5A/BrettsSandbox/paper/data/VariantSCN5A-new.db")
alltables = dbListTables(con)
my.data <- dbReadTable(con, 'VariantSCN5A')
my.data[my.data=='NA'] <- NA
d<-my.data
dbDisconnect(con)

d$resnum<-as.integer(d$resnum)
d$gnomAD[is.na(d$gnomAD)] <- 0
d$gnomAD<-as.numeric(d$gnomAD)
d$ipeak<-100*as.numeric(d$ipeak)
d$vhalfact<-as.numeric(d$vhalfact)
d$vhalfinact<-as.numeric(d$vhalfinact)
d$recovfrominact<-log10(100*as.numeric(d$recovfrominact))
d$ilate[as.numeric(d$ilate)==0]<-NA
d$ilate_norm<-log10(d$ipeak*as.numeric(d$ilate)+0.00001)
d$ilate<-log10(100*as.numeric(d$ilate)+0.00001)
d$total_carriers<-d$lqt3+d$brs1+d$unaff+d$gnomAD
d$weight = 1-1/(0.1+d$total_carriers) #weights
d$weightsMilder = 1-1/(1+d$total_carriers) #weights
d$noweights = rep(1,length(d$total_carriers))


servers<-read.csv("/Users/B/Dropbox/SCN5A/BrettsSandbox/paper/data/annotated_variants-trim.txt", sep = "\t")
provean <-read.csv("/Users/B/Dropbox/SCN5A/BrettsSandbox/paper/data/provean.txt", sep = "\t")
pph2 <-read.csv("/Users/B/Dropbox/SCN5A/BrettsSandbox/paper/data/pph2-short.txt", sep = "\t")
sift <-read.csv("/Users/B/Dropbox/SCN5A/BrettsSandbox/paper/data/SIFT.txt", sep = "\t")
d <- merge(d, servers, all = TRUE)
d <- merge(d, provean, all = TRUE)
d <- merge(d, sift, all = TRUE)
d <- merge(d, pph2, all = TRUE)
d<-d[!is.na(d$var), ]

d$eaRate<-as.numeric(d$eaRate)


# Adding in penetrance variables
abrs0=0.32 
alqt0=0.11   
beta0=1 
d$LQT_penetranceBayesian<-(d$lqt3+alqt0)/(d$total_carriers+beta0+alqt0)
d$BrS_penetranceBayesian<-(d$brs1+abrs0)/(d$total_carriers+1+alqt0)
d$all_penetranceBayesian<-(d$brs1+abrs0+d$lqt3+alqt0)/(d$total_carriers+beta0+alqt0+abrs0)
e<-d

```


#plot probability distributions for BrS1/LQT3 included in main text

G1748D, D1790G, R965C, L1501V, I1660V, R1644C, T1304M, and E1784K

```{r sample distributions, fig.width=4}
uqs<-c("G1748D","D1790G","R965C","L1501V","I1660V","R1644C","T1304M","E1784K")
vois<-c(0)
for(v in 1:length(uqs)){
  vois[v]<-match(uqs[v],d$var)
}
x <- seq(0,.99,0.01)
par(lwd = 2)#, mfrow=c(1,3)) 
for(i in vois){
  mb <- max(dbeta(x[2:(length(x)-1)],abrs0+d$brs1[i], beta0+(d$total_carriers[i]-d$brs1[i])))
  ml <- max(dbeta(x[2:(length(x)-1)],alqt0+d$lqt3[i], beta0+(d$total_carriers[i]-d$lqt3[i])))
  mm <- max(ml,mb)
  curve(dbeta(x, alqt0+d$lqt3[i], beta0+(d$total_carriers[i]-d$lqt3[i])), 
        ylab="Probability", 
        xlab = "Penetrance", add = FALSE, lty = 1, ylim = c(0,mm))
  curve(dbeta(x, abrs0+d$brs1[i], beta0+(d$total_carriers[i]-d$brs1[i])), 
        add = TRUE, lty=2)
  text(0.5,mm*0.92,paste("BrS1: ", d$brs1[i],
                         " LQT3: ", d$lqt3[i],
                         " Total: ", d$total_carriers[i]))
}



```

# Plot integrated probability distributions for variants classifiable as (likely) pathogenic, VUS, or (likley) benign

These classifications were made based on the ACMG guidelines from the following website: 
http://www.medschool.umaryland.edu/Genetic_variant_Interpretation_Tool1.html/

variables used to classfy:
PS3/BS3 in vitro functional studies COMPROMIZED/NORMAL
PS4 enriched in affected population
BS1/PM2 TOO HIGH/LOW in gnomad
PP2 Missense var in gene with low rate of benign missense vars
PP3/BP4 Multiple lines of computational evivence FOR/AGAINST pathogenicity

Set disease rate and enrichment threasholds

```{r setparms}
dr = 15 # disease rate threashold in gnomad. I'm just going to select one, 8 for BrS1 and 7 for LQT3 
et = 0.2 # enrichment threashold. The fraction of carriers who need to present
         # with a disease in order to use PS4. This is subjective
```

## Likely pathogenic (III) 
given PS4, PP2, PM2

```{r del}
d <- e[e$Prediction=="*DAMAGING" & 
         (e$prediction=="possiblydamaging" | e$prediction=="probablydamaging"),] # PP3
d <- d[d$gnomAD<=dr | is.na(d$gnomAD),] # PM2
d <- d[(d$lqt3+d$brs1)/(d$total_carriers)>et,] # PS4
d <- d[!is.na(d$var),]

del <- matrix(0, nrow=length(x), ncol=2)
new.mat.d <- matrix(nrow=length(d$var), ncol=4)
for(i in 1:length(d$var)){
  new.mat.d[i,] <- c(as.character(d$var[i]),
                     d$lqt3[i],d$brs1[i],d$total_carriers[i])
  
  for(j in 1:length(x)){
    del[j,] <- c(x[j], del[j,2]+integrate(dbeta, shape1 = abrs0+alqt0+as.numeric(new.mat.d[i,2])+as.numeric(new.mat.d[i,3]),
                                          shape2 = (beta0+as.numeric(new.mat.d[i,4])-
                                          as.numeric(new.mat.d[i,2])-as.numeric(new.mat.d[i,3])), x[j], x[j]+0.01)$val
                 )
  }
}
colnames(new.mat.d) <- c("variant", "lqtpos", "brspos", "vartot")
head(new.mat.d)
```

## Likely benign (I) 
given PP2, BP4, BS1 

```{r benign}
b <- e[e$Prediction!="*DAMAGING" & 
         (e$prediction!="possiblydamaging" & e$prediction!="probablydamaging"),] # BP4
b <- b[b$gnomAD>=dr | is.na(b$gnomAD),] # BS1
b <- b[!is.na(b$var),]

ben <- matrix(0, nrow=length(x), ncol=2)
new.mat.b <- matrix(nrow=length(b$var), ncol=4)
for(i in 1:length(b$var)){
  new.mat.b[i,] <- c(as.character(b$var[i]),b$lqt3[i], b$brs1[i], b$total_carriers[i])
  for(j in 1:length(x)){
    ben[j,] <- c(x[j], ben[j,2]+integrate(dbeta, shape1 = abrs0+alqt0+as.numeric(new.mat.b[i,2])+as.numeric(new.mat.b[i,3]),
                                          shape2 = (beta0+as.numeric(new.mat.b[i,4])-
                                          as.numeric(new.mat.b[i,2])-as.numeric(new.mat.b[i,3])), x[j], x[j]+0.01)$val)
  }
}
colnames(new.mat.b) <- c("variant", "lqtpos", "brspos", "vartot")
head(new.mat.b)
```

## VUS 
given PM2, BP4, PP2

```{r VUS}
v <- e[e$Prediction!="*DAMAGING" & 
         (e$prediction!="possiblydamaging" & e$prediction!="probablydamaging"),] # BP4
v <- v[v$gnomAD<=dr | is.na(v$gnomAD),] # PM2
v <- v[!is.na(v$var),]

# vUSs 
vus <- matrix(0, nrow=length(x), ncol=2)
new.mat.v <- matrix(nrow=length(v$var), ncol=4)
for(i in 1:length(v$var)){
  new.mat.v[i,] <- c(as.character(v$var[i]),v$lqt3[i], v$brs1[i], v$total_carriers[i])
  for(j in 1:length(x)){
    vus[j,] <- c(x[j], vus[j,2]+integrate(dbeta, shape1 = abrs0+alqt0+as.numeric(new.mat.v[i,2])+as.numeric(new.mat.v[i,3]),
                                          shape2 = (beta0+as.numeric(new.mat.v[i,4])-
                                          as.numeric(new.mat.v[i,2])-as.numeric(new.mat.v[i,3])), x[j], x[j]+0.01)$val
                 )
  }
}
colnames(new.mat.v) <- c("variant", "lqtpos", "brspos", "vartot")
head(new.mat.v)
```

## Plot probability distribution

```{r plots}

scaled<-length(x)/(length(d$var))
scaleb<-length(x)/(length(b$var))
scalev<-length(x)/(length(v$var))

#plot all variants classified as benign
plot(ben[,1],ben[,2]*scaleb,ylim = c(0,2.5),type = "l", ylab = "Probability integrated over all Benign Variants", xlab = "Fraction of Carriers Presenting with BrS1 or LQT3", lwd = 3)

#plot all variants classified as VUS
lines(vus[,1],vus[,2]*scalev,ylim = c(0,2.5),type = "l", ylab = "Probability integrated over all VUS Variants", xlab = "Fraction of Carriers Presenting with BrS1 or LQT3", lwd = 3)

#plot all variants classified as pathogenic
lines(del[,1],del[,2]*scaled,ylim = c(0,2.5),type = "l", ylab = "Probability integrated over all Pathogenic Variants", xlab = "Fraction of Carriers Presenting with BrS1 or LQT3", lwd = 3)
```

## Calculate the probability the "average/typical" variant has a posterior mean penetrance fraction of carriers presenting with either Brs1 or LQT3 > 20%

```{r prob20p}
# calculate integrated probability of penetrance > 20% (1:5). 
delh<-0
for (j in 1:(length(x)-length(x)/5)){
  delh <- delh+del[j+length(x)/5,2]*.01*scaled
}
vush<-0
for (j in 1:(length(x)-length(x)/5)){
  vush <- vush+vus[j+length(x)/5,2]*.01*scalev
}
benh<-0
for (j in 1:(length(x)-length(x)/5)){
  benh <- benh+ben[j+length(x)/5,2]*.01*scaleb
}

print("probability a variant classified as (likely) pathogenic has a penetrance > 20%")
delh
print("probability a variant classified as VUS has a penetrance > 20%")
vush
print("probability a variant classified as (likely) benign has a penetrance > 20%")
benh

#run the following in console:\n
#pdf("~/../Dropbox/Andrew-Brett/scn5a_annotation/paper/images/class_<number>.pdf")\n
#<all the desired plots written below>\n
#dev.off()\n
```

# Sensitivity Analysis
Solid lines are the prior, dashed lines are posterior if one affected carrier is observed

## Empirical Bayes (used in manuscript)

```{r sensitivity, fig.width=7, fig.height=7}

# Penetrance calculation used (Manuscript Version)
abrs0=0.32 
alqt0=0.11 
beta0=1 
curve(dbeta(x, abrs0, beta0), ylab="Probability", xlab="Penetrance", lwd = 3)
curve(dbeta(x, abrs0+1, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, lty = 2, add = TRUE)
```

## Uninformative Prior

```{r uninformative, fig.width=7, fig.height=7}
# Changing penetrance calculation to uninformative prior
abrs0=1
alqt0=1
beta0=1 
curve(dbeta(x, abrs0+1, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, lty = 2)
curve(dbeta(x, abrs0, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, add = TRUE)
```

## Optimistic Prior

```{r optimistic, fig.width=7, fig.height=7}
# Changing penetrance calculation to optimistic (no affected carriers)
abrs0=0.01
alqt0=0.01
beta0=1 
curve(dbeta(x, abrs0+1, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, lty = 2)
curve(dbeta(x, abrs0, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, add = TRUE)
```

## Pessimistic Prior

```{r pessimistic, fig.width=7, fig.height=7}
# Changing penetrance calculation to pessimistic (one affected carrier)
abrs0=1
alqt0=1
beta0=0.01 
curve(dbeta(x, abrs0, beta0), ylab="Probability", xlab="Penetrance", lwd = 3)
curve(dbeta(x, abrs0+1, beta0), ylab="Probability", xlab="Penetrance", lwd = 3, lty = 2, add = TRUE)
```